### 2016-10-04 HRI/Parts/CaseStudy.tex
Changes in the HRI article

The case study was done at Researcher's Night 2015 with two main objectives (i) cross-validate the findings obtained from the experiment, and (ii) use the Emotional Enrichment System to verify whether the participants would prefer scenes when the robot expresses emotions or rather moves without any emotion expression.

\subsection{Design and Setup}

This case study uses the results obtained in the experiment and was designed to have two parts (i) emotion expression through changes in linear velocity, angular velocity, oscillation angle, orientation and direction.  
(ii) Presentation of a small scene to verify whether the participants would prefer scenes when the robot shows emotions or not. the Emotional Enrichment System was used in both cases (with and without emotions). Two web-cams and eight Alvar tags were added to use Kalman filter to improve robots localization in the stage. The detection of the AR tags is done through the use of the ROS package ar\_track\_alvar~\cite{artag2015}. The distribution of the web-cams and the tags are depicted in Figure~\ref{fig:setup_fourth}.

@@

\subsubsection{Emotion Description}

The parameters selected for each of the four emotions (\textit{Anger}, \textit{Happiness}, \textit{Sadness} and \textit{Fear}) are shown in Table~\ref{table:selected_fourth}. The main considerations to select the two implementations for each emotion were: (i) the linear velocity should be greater than $0$. In other words the robot should show some linear displacement. And (ii) it should be in the top 10 list of the emotion obtained in the experiment.

@@

\subsubsection{Scene}
%TODO change this part of the text and give some information about this, and  motivate its use, which bassically is to verify the emotional enrichment system
As it was already mentioned, actors should adapt to different circumstances during the performance. Following this approach, the stage was discretized in 9x9 matrix as is shown in Figure~\ref{fig:stage_division}. Thus, the movements of the robot are given in terms of the matrix positions. This allows the adaptation to different stage dimensions because the robot's final position is calculated by the Emotional Enrichment System during execution taken under consideration the stage dimensions. For instance, during the scene's preparation in the laboratory the stage was 3 meters per 3 meters, but in the final presentation the stage was 2.5 meters per 2.5 meters.

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{./Images/FourthCaseScene.png} 
	\caption{Stage discretization  used for the small scene. The blue squares correspond to the each zone, while the numbers correspond to the ID given to each zone.}
	\label{fig:stage_division}
\end{figure} 
%TODO Change this paragraph and motivate the reason this is important and instead of puting the figures, write the references to those figures
The scene's description is the following: the robot starts in the middle of the stage to move to the upstage right (Figure~\ref{fig:StageDirections}), close to the right wing. Then, the robot moves to upstage right center and rotates by $\pi/2$ left (Figure~\ref{fig:BodyPosition}). Next the robot moves to the right center to then go to the center. When it arrives there, it turns full back and move backwards to downstage center with a full front orientation. There, it turns full back to move to center. Finally the robot turns to profile right and it does a step back; then it goes to the upstage center and then upstage right. The sequence of movements programmed to the robot are depicted in Figure~\ref{fig:movement}.
\begin{figure*}
	\centering
	\includegraphics[width=0.85\textwidth]{./Images/fourthCaseSceneD.png} 
	\caption{Sequence of movements done by the robot. The red arrows show the trajectory done by the robot, while the numbers show the order among the movements. a) The first ten movements b) The last five movements }
	\label{fig:movement}
\end{figure*}

The relation between emotion and movement is as follow: movements one, two, three, four and five are expressed without any emotion. Movements six, seven, eight, nine and ten show fear. Movement eleven depicts happiness, and the remaining movements depict sadness. The two scenes are executed by the Emotional Enrichment System using the same ''script'' and the emotion selection is done manually via graphical interface (Figure~\ref{fig:graphical_interface}).The ''script'' were written in JSON using the language described in the Section~\ref{}.

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{./Images/InterfaceExperiment.png} 
	\caption{Graphical interface used to communicate with the Emotional Enrichment System. a) It is the interface used to send the actions sequence. b) It is the interface used to send a emotion and its intensity.}
	\label{fig:graphical_interface}
\end{figure}

>>>>

The case study was done at Researcher's Night 2015 with two main objectives (i) cross-validate the findings obtained from the experiment, and (ii) use the Emotional Enrichment System to verify whether the participants would prefer scenes when the robot expresses emotions or rather moves without any emotion expression. The Emotional Enrichment System was used in both cases. To reduce scene variability was used two web-cams and eight Alvar tags to informed a Kalman filter to improve robots localization in the stage. The detection of the AR tags was done through the use of the ROS package ar\_track\_alvar~\cite{artag2015}. The distribution of the web-cams and the tags are depicted in Figure~\ref{fig:setup_fourth}.

@@

\subsection{Emotion Description}

The parameters selected for each of four emotions (\textit{Anger}, \textit{Happiness}, \textit{Sadness} and \textit{Fear}) are shown in Table~\ref{table:selected_fourth}. The main considerations to select the two implementations for each emotion were: (i) the linear velocity should be greater than $0$. In other words the robot should show some linear displacement. And (ii) it should be in the top 10 list of the emotion obtained in the experiment.

@@

\subsection{Scene}

Stage discretization was used to give zones of movements instead of absolute positions. This idea was brought from human theatrical actors, who prepared their movements based on zones on the stage~\cite{wilson2009theatre}. This allows them to adapt their position based on other actors. Therefore, the stage was discretized in 9x9 matrix as is shown in Figure~\ref{fig:stage_division}. Robot's movements are given in terms of the matrix positions to the Emotional Enrichment System. This allows the adaptation to different stage dimensions because robot's final position is calculated by the Emotional Enrichment System during execution, which takes under consideration the stage dimensions given. For instance, during the scene's preparation in the laboratory the stage was 3 meters per 3 meters, but in the final presentation the stage was 2.5 meters per 2.5 meters.

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{./Images/FourthCaseScene.png} 
	\caption{Stage discretization  used for the small scene. The blue squares correspond to the each zone, while the numbers correspond to the ID given to each zone.}
	\label{fig:stage_division}
\end{figure}

The scene's description is the following: the robot starts in the middle of the stage to move to the upstage right (See~\cite{Musical}), close to the right wing. Then, the robot moves to upstage right center and rotates by $\pi/2$ left (See~\cite{Artopia}). Next the robot moves to the right center to then go to the center. When it arrives there, it turns full back and move backwards to downstage center with a full front orientation. There, it turns full back to move to center. Finally the robot turns to profile right and it does a step back; then it goes to the upstage center and then upstage right. The sequence of movements programmed to the robot are depicted in Figure~\ref{fig:movement}.
\begin{figure*}
	\centering
	\includegraphics[width=0.95\textwidth]{./Images/fourthCaseSceneD.png} 
	\caption{Sequence of movements done by the robot. The red arrows show the trajectory done by the robot, while the numbers show the order among the movements. a) The first ten movements b) The last five movements }
	\label{fig:movement}
\end{figure*}

The relation between emotion and movement is as follow: movements one, two, three, four and five are expressed without any emotion. Movements six, seven, eight, nine and ten show fear. Movement eleven depicts happiness, and the remaining movements depict sadness. The two scenes are executed by the Emotional Enrichment System using the same ''script'' and the emotion selection is done manually via graphical interface (Figure~\ref{fig:graphical_interface}).

\begin{figure}
	\centering
	\includegraphics[width=0.48\textwidth]{./Images/InterfaceExperiment.png} 
	\caption{Graphical interface used to communicate with the Emotional Enrichment System. a) It is the interface used to send the actions sequence. b) It is the interface used to send a emotion and its intensity.}
	\label{fig:graphical_interface}
\end{figure}

### 2016-10-04 HRI/Parts/Conclusions.tex
Changes in the HRI article

The case study presented in this chapter was done to cross validate the findings in the experiment. For each one of the four emotions studied in the experiment were selected two set of parameters. The results show that both implementations of happiness were confused with anger and excitement, while one implementation of anger was just confused with excitement. Both implementations of sadness were confused with tenderness and fear. Both implementations of fear had a recognition rate over 50\%. Additionally to the cross validation, it was done a small scene to check if people have preference to scenes where emotional movements are presented or not. The results show that people prefer scenes with emotional movements.

>>>>

The case study presented in this paper was done to cross validate the findings in the experiment and verify whether the participants would prefer scenes when the robot expresses emotions or rather moves without any emotion expression. For each one of four emotions (i.e. Anger, Happiness, Sadness and Fear) studied in the experiment were selected a two set of parameters. These parameters were described in files and used as input to the Enrichment Emotional System, which was in charge to combine actions with emotions. The results show that both implementations of happiness were confused with anger and excitement, while one implementation of anger was just confused with excitement. Both implementations of sadness were confused with tenderness and fear. Both implementations of fear had a recognition rate over 50\%. Scene's results show that people prefer scenes with emotional movements and there is not any difference in gender.

Additionally to the results already mentioned, there are words that could bias participants' perception. For instance happiness and anger were considered as excitement. This misinterpretation should not be a surprise given the fact that there is not a unique definition of emotion~\cite{Plutchik2001,cacioppo2000handbook}, and each person would interpret a situation differently, so they will give a different label to the presented movement. Moreover a misinterpretation of Happiness and Anger could suggest that additional features (e.g. trajectory or shape) should be added to increase differentiation between these two emotions. For example, Venture and collaborators~\cite{Venture2014} had found out that in human bodies the recognition rate of anger and fear are increased when torso and head are downwards. On the other hand, they found that happiness perception is increased when the torso and head are move upwards. This example could bring some insight to possible body changes that could occur in non-human like bodies, but it should tested in this kind of platforms to confirm if the same impact is reached.

### 2016-10-04 HRI/Parts/RelatedWork.tex
Changes in the HRI article

The direct consequence of the abundance of works in face elicitation in humans is the amount of works done in Human-Robot Interaction (HRI) focused also on faces. One of the most well-known expressive robots is Kismet~\cite{Breazeal2002}, a robotic face able to interact with people and to show emotions. The face had enough degrees of freedom to portray the basic emotions suggested by Ekman~\cite{Ekman2004} (\textit{happiness}, \textit{surprise}, \textit{anger}, \textit{disgust}, \textit{fear}, and \textit{sadness}) plus \textit{interest}.  Despite the complex system behind Kismet, the emotion's projection evaluation was done using videos with a very limited number of participants. In this experiment, each participant was asked to look at seven videos, one for each implemented emotion. After a video was projected, each participant selected the emotion from a list containing  the seven implemented emotions, only. The results showed high percentage of recognition (over 57\%).

%%%%%%%%%%%%%%%%%%%%%%%%
Saerbeck and Christoph~\cite{Saerbeck2010} analyzed the relationship between robots' motion characteristics and perceived affect. They first did a literature review to select characteristics that could be used to show affection. From their literature review, they decided to use acceleration and curvature on robot's trajectory. 
Using these two characteristics, they did an experiment with eighteen participants. They used two different platforms (iCat and Roomba) to verify if the embodiment had any impact on affection determination. Each participant was exposed to both embodiments and all possible variable combinations. To assess participants' perception, they used PANAS~\cite{WatsonClarkTellegen88} and Self-Assessment Manikin (SAM)~\cite{Lang2008}. Their results show that there was no significant difference between the two embodiments used and that participants were able to assign different emotions to the movement patterns shown to them. 
More importantly, they found out that acceleration is correlated with the perceived arousal but not with valence.

%%%%%%%%%%%%%%%%
Ca\~namero and collaborators~\cite{Canamero2010,Beck2010} studied the perception of key body poses designed to show emotions using a NAO~\cite{NAO2013}. They suggested that the techniques used to convey emotions in virtual characters could not be used in robots, due to the fact that virtual characters have no physical constraints, while robots are constrained by their physical capabilities~\cite{Saerbeck2007, Canamero2010}. They proposed a set of poses that could be used to express emotions with NAO, paying a particular attention to the contribution of the head. Their results showed a recognition rate of 88\%, 85\%, 92\%, 88\%, 73\%, and 73\% for anger, sadness, fear, pride, happiness and excitement respectively. The major finding from their experiments is that moving the head up improved the identification of pride, happiness, and excitement. While moving the head down improved the identification of anger and sadness.

@@

%%%%%%%%%%%%%%%%%%%%%%%
Using the platform WABIAN-2R, Destephe and collaborators~\cite{Destephe2013b} studied the attribution of emotion to a robot's gait. To obtain the robot's movements, the researchers asked two professional actors to walk in a room conveying anger, happiness, sadness, and fear with different intensities (low, regular, normal, high, and exaggerated). All the actors' walking were recorded using a Cortex motion capture system. These data were later reported to robot's embodiment. Each subject was exposed to a series of videos. After each video, the subject was asked to select one emotion from a listand to state the intensity of the selected emotion. The videos showed were not made with the real robot, but with a virtual model of the robot. Their results showed that sadness was recognized 73.81\% (average) of the times with an intensity of 21.43\%, happiness was recognized 66.67\% (average) of the times with an intensity of 30.95\%, anger was recognized 61.9\% of the times with an intensity of 26.19\%, and fear was recognized 83.33\% (average) of the times with an intensity of 28.58\%. These results show that people could reasonably well recognize emotions from the robot's gait.

%%%%%%%%%%%%%%%%% %
Brown and Howard focused their interest on head and arm movements using the DARwIn-OP platform~\cite{Brown2014}. Their hypothesis was that using some principles, that they determined as important, it is possible to express happiness and sadness in a way that people could recognize. These principles establish that to show happiness is necessary to move robot's head up, arms up, and it should be done fast. While to express sadness the head goes down, arms down, and it should be done slow. To test their hypothesis they conducted an experiment with thirteen participants. Each participant was exposed to fifteen sequences of poses and he/she had to answer for each sequence a five-point Likert scale (very-sad to very happy, with neutral in the middle). They obtained a 95\%, 59\%, and 94\% of accuracy for happiness, neutral and sadness respectively.

@@

%%%%%%%%%%%%%%%%%%%
To improve the coordination between humans and robots Novikova and Watss~\cite{Novika2015} proposed the use of emotion. They did studies using a own built platform called E4 to verify whether people could detect emotions from a non-human like platform. The E4 platform was constructed using a Lego Mindstorms NTX and it was based on a Phobot robot's design
. In all their experiments they used six emotional expressions (scared, surprised, excited, angry, happy, sad) and neutral. The emotion list used in their experiments was balanced using the 2-D circumplex model (arousal and valence), selecting three from each quadrant for a total of twelve emotions, plus the options of ``other'' and ``don't know''. They obtained a recognition rate of 52\%, 42\%, 41\%, 36\% and 15\% for surprise, fear, sadness, happiness, and anger respectively.

>>>>

A direct consequence of the abundance of works in face elicitation in humans is the amount of works done in Human-Robot Interaction (HRI) focused also on faces. One of the most well-known expressive robots is Kismet~\cite{Breazeal2002}, a robotic face able to interact with people and to show emotions. The face had enough degrees of freedom to portray the basic emotions suggested by Ekman~\cite{Ekman2004} (\textit{happiness}, \textit{surprise}, \textit{anger}, \textit{disgust}, \textit{fear}, and \textit{sadness}) plus \textit{interest}.  Despite the complex system behind Kismet, the emotion's projection evaluation was done using videos with a very limited number of participants. In this experiment, each participant was asked to look at seven videos, one for each implemented emotion. After a video was projected, each participant selected a emotion from a list, which contained seven implemented emotions. The results showed high percentage of recognition (over 57\%).

%%%%%%%%%%%%%%%%%%%%%%%%
Saerbeck and Christoph~\cite{Saerbeck2010} analyzed the relationship between robots' motion characteristics and perceived affect. They first did a literature review to select characteristics that could be used to show affection. From their review, they decided to use acceleration and curvature on robot's trajectory. 
Using these two characteristics, they did an experiment with eighteen participants. They used two different platforms (iCat and Roomba) to verify if the embodiment had any impact on affection determination. Each participant was exposed to both embodiments and all possible variable combinations. To assess participants' perception, they used PANAS~\cite{WatsonClarkTellegen88} and Self-Assessment Manikin (SAM)~\cite{Lang2008}. Their results show that there was no significant difference between the two embodiments used and that participants were able to assign different emotions to the movement patterns shown to them. 
More importantly, they found out that acceleration is correlated with the perceived arousal but not with valence.

%%%%%%%%%%%%%%%%
Ca\~namero and collaborators~\cite{Canamero2010,Beck2010} studied the perception of key body poses designed to show emotions using a NAO~\cite{NAO2013}. They suggested that techniques used to convey emotions in virtual characters could not be used in robots, due to the fact that virtual characters have no physical constraints, while robots are constrained by their physical capabilities~\cite{Saerbeck2007, Canamero2010}. They proposed a set of poses that could be used to express emotions with a NAO, paying a particular attention to head's contribution. Their results showed a recognition rate of 88\%, 85\%, 92\%, 88\%, 73\%, and 73\% for anger, sadness, fear, pride, happiness and excitement respectively. The major finding from their experiments is that moving the head up improved the identification of pride, happiness, and excitement. While moving the head down improved the identification of anger and sadness.

@@

%%%%%%%%%%%%%%%%%%%%%%%
Using the platform WABIAN-2R, Destephe and collaborators~\cite{Destephe2013b} studied the attribution of emotion to a robot's gait. To obtain  robot's movements, the researchers asked two professional actors to walk in a room conveying anger, happiness, sadness, and fear with different intensities (low, regular, normal, high, and exaggerated). All the actors' walking were recorded using a Cortex motion capture system. These data were later reported to robot's embodiment. Each subject was exposed to a series of videos. After each video, the subject was asked to select one emotion from a list and state the intensity of the selected emotion. The videos showed were not made with the real robot, but with a virtual model of the robot. Their results showed that sadness was recognized 73.81\% (average) of the times with an intensity of 21.43\%, happiness was recognized 66.67\% (average) of the times with an intensity of 30.95\%, anger was recognized 61.9\% of the times with an intensity of 26.19\%, and fear was recognized 83.33\% (average) of the times with an intensity of 28.58\%. These results show that people could reasonably well recognize emotions from the robot's gait.

%%%%%%%%%%%%%%%%% %
Brown and Howard focused their interest on head and arm movements using a DARwIn-OP platform~\cite{Brown2014}. Their hypothesis was that using some principles, that they determined as important, it is possible to express happiness and sadness in a way that people could recognize. These principles establish that to show happiness is necessary to move robot's head up, arms up, and it should be done fast. While to express sadness the head goes down, arms down, and it should be done slow. To test their hypothesis they conducted an experiment with thirteen participants. Each participant was exposed to fifteen sequences of poses and he/she had to answer for each sequence a five-point Likert scale (very-sad to very happy, with neutral in the middle). They obtained a 95\%, 59\%, and 94\% of accuracy for happiness, neutral and sadness respectively.

@@

%%%%%%%%%%%%%%%%%%%
To improve the coordination between humans and robots Novikova and Watss~\cite{Novika2015} proposed the use of emotion. They did studies using a own built platform called E4 to verify whether people could detect emotions from a non-human like platform. The E4 platform was constructed using a Lego Mindstorms NTX and it was based on a Phobot robot's design. In all their experiments they used six emotional expressions (scared, surprised, excited, angry, happy, sad) and neutral. The emotion list used in their experiments was balanced using the 2-D circumplex model (arousal and valence), selecting three from each quadrant for a total of twelve emotions, plus the options of ``other'' and ``don't know''. They obtained a recognition rate of 52\%, 42\%, 41\%, 36\% and 15\% for surprise, fear, sadness, happiness, and anger respectively.
